Sampling Strategies in Decisions from Experience

### 2022-03-11 manuscript.Rmd
Add appendix with anlyses of comprehensive simulation

bibliography      : ["references_sampling-in-dfe.bib"]
csl               : apa.csl

@@

classoption       : "man"
documentclass     : "apa7"
output            : papaja::apa6_pdf

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, fig.pos = "t", fig.align = "l")
```

```{r packages}
# load required packages
pacman::p_load(papaja,
               readr,
               tidyr, 
               dplyr,
               ggplot2,
               viridis,
               ggpubr,
               latex2exp)
```

@@

Similar to other behavioral phenomena that can be explained by the mind being sensitive to the composition of the samples on which it operates [see @fiedlerBewareSamplesCognitiveecological2000], the as-if underweighting of rare outcomes in DfE can be explained by peoples' tendency to rely on small samples [@hertwigDecisionsExperienceEffect2004; @rakowBiasedSamplesNot2008; @wulffMetaanalyticReviewTwo2018].
That is, following from the laws of large numbers [LLN, see, e.g., @kolmogorovFoundationsTheoryProbability1950], the relative frequency with which a prospect's outcome occurs in a sample of infinite size converges almost surely to its objective probability.
In such a case, then, the binomial distribution associated with each prospect's outcome can be approximated by a symmetric normal distribution [cf. @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015].
However, for small samples, the binomial distribution associated with small-probability outcomes is positive skewed, therefore causing the relative frequencies with which these outcomes occur in the sample to be smaller rather than larger than their latent objective probabilities.
Thus, small-probability outcomes tend to be underrepresented in small samples and might not occur at all.
As a consequence, were the mind to use the sampled relative frequencies of outcomes to infer their latent objective probabilities and to follow the principle of mathematical expectation, the reliance on small samples could explain the as-if underweighting of rare outcomes.

Aside this computational-level explanation, different algorithmic approaches linking the sampling process to the choices in DfE can explain the as-if underweighting of rare outcomes by the reliance on small samples.
That is, in one paradigm used to study DfE, namely, the sampling paradigm [@weberPredictingRiskSensitivity2004], before making a single choice between prospects with latent properties, people can, as much as they want, sequentially sample the prospects' outcomes according to their objective probabilities.
To couple the sampling with the decision-making process, sequential sampling models [e.g., @bhatiaSequentialSamplingParadoxes2014; @busemeyerDecisionFieldTheory1993; @markantModelingChoiceSearch2015; @ratcliffComparisonSequentialSampling2004] propose that evidence is accumulated over time.
The core assumption shared by these models is that the sampled outcomes are sequentially integrated into one or more dynamic decision variables, and that a choice is made once a decision variable exceeds a threshold in favor for one of the prospects.
Moreover, these models use differences in thresholds to explain speed-accuracy trade-offs in decision making, with higher thresholds requiring more evidence in favor of a prospect to be sampled.
Hence, since the overall number of sampled outcomes decreases with thresholds, the adoption of low thresholds can be used to explain an as-if underweighting of rare outcomes [see @markantModelingChoiceSearch2015].

Yet, in another approach, @hillsInformationSearchDecisions2010 propose a dependency between sampling patterns and decision strategies.
Considering the choice between two prospects, this approach distinguishes sampling patterns along the frequency of switches between the prospects during the sampling phase and assumes that only outcomes that were sampled from the same prospect between switches are integrated into the same sample. 
Then, it is assumed that each time after a new sample for both prospects was generated, their means are compared, and the choice is determined by the number of won comparisons.^[<!--Open footnote-->Note that @hillsInformationSearchDecisions2010 describe only the two paradigmatic cases, where prospects are either switched just once (comprehensive sampling) or after each sampled outcome (piecewise sampling). Accordingly, in the former case choices depend on a single comparison between the means of samples that contain all outcomes sampled for the respective prospect (summary strategy); in the latter case, however, choices depend on multiple comparisons between samples that contain only one sampled outcome (round-wise strategy). Nevertheless, Hills and Hertwig [-@hillsInformationSearchDecisions2010, p. 1788] acknowledge that "many [sampling] strategies will fall on the continuum between [these cases]", which is why the general description is used above.<!--Close footnote-->]
Since each comparison is weighted equally, such a link between sampling patterns and decision strategies can explain an as-if underweighting of rare outcomes despite a large overall number of sampled outcomes.
Indeed, some studies have found the description-experience gap to appear, even if the overall number of sampled outcomes is large and the relative frequencies across all sampled outcomes closely resemble the latent objective probabilities [@hauDescriptionexperienceGapRisky2008; @hauDecisionsExperienceStatistical2010; @ungemachAreProbabilitiesOverweighted2009; see also @wulffMetaanalyticReviewTwo2018].
More specifically, while the probability of a rare outcome to be sampled increases with the overall number of sampled outcomes, an increasing switching frequency causes the rare outcome not to be considered in an increasing number of comparisons.

@@

The paper proceeds as follows:
The first section provides a formal definition of prospects and reviews cumulative prospect theory [CPT, @tverskyAdvancesProspectTheory1992; @wakkerAxiomatizationCumulativeProspect1993] as a data model that can be used to describe choices between prospects in terms of deviations from the principle of mathematical expectation.
In the section that follows, a computational model is proposed that integrates the assumption of evidence accumulation over time and of a link between sampling patterns and decision strategies.

The paper closes with a summary and a conclusion.

@@

\noindent
Given that the value of each cumulative decision weight depends on the two transformed cumulative probabilities in (12)---i.e., the probability of obtaining a positive (negative) outcome equal to or greater (smaller) than a respective outcome $x$, and the probability of obtaining a strictly greater (smaller) outcome---the remaining consequences for the weighting pattern are reviewed by considering the actual weighting function.
Therefore, Figure\ \@ref(fig:w-shapes) illustrates some of the possible graphical shapes of the two-parameter weighting function of @goldsteinExpressionTheoryPreference1987, which, however, is just one of several parameterizations that have been proposed [e.g., @prelecProbabilityWeightingFunction1998; @tverskyAdvancesProspectTheory1992; see @stottCumulativeProspectTheory2006, for an overview].
Now, let each $p$ on the abscissa be one of the four cumulative probabilities from (12). 
Then each graph in Figure\ \@ref(fig:w-shapes) displays the graphical shape of the weighting function
$$
\begin{aligned}
  w : [0,1] &\to [0,1] \\
  p &\mapsto \frac{\delta \times p^{\gamma}}
  {\delta \times p^{\gamma} + (1-p)^{\gamma}}
  \; ,
\end{aligned}
\tag{11}
$$
for the respective values of the parameters $\gamma \in [0,2]$ and $\delta > 0$.
Evidently, both parameters have distinct effects on the graphs' shape, with $\gamma$ affecting the curvature and $\delta$ the elevation.
As a consequence from (12) then, each combination of parameters implies a particular weighting pattern, with some of them being similar and others being rather distinct.

(ref:weighting) Possible Graphical Shapes of Goldstein and Einhorn's -@goldsteinExpressionTheoryPreference1987 Weighting Function

```{r w-shapes, fig.cap="(ref:weighting)", fig.height=7, fig.width=10}

# prepare vector with cumulative probabilities
p <- tibble(p = seq(0, 1, .01))

# compute weighting function for different combinations of gamma and delta 
wf <- p %>% expand_grid(gamma = seq(.1, 2, .1), 
                        delta = c(.1, .5, 1, 2, 5, 10)) %>%
  mutate(wp = (delta*(p^gamma))/((delta*p^gamma)+(1-p)^gamma))

# labeller function for LaTeX math expressions in facet labels 
labeller_delta <- function(string) {
  TeX(paste("$\\delta=$", string, sep = ""))  
}

# plot
ggplot(wf, aes(p, wp, group = gamma)) +
  facet_wrap(~delta, labeller = as_labeller(labeller_delta, default = label_parsed)) + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x = expression(p),
       y = expression(w(p)), 
       color = expression(gamma)) + 
  theme_apa() + 
  geom_line(aes(color = gamma)) +
  scale_color_viridis(option = "plasma") + 
  geom_abline(intercept = 0, slope = 1, color = "gray", linetype = "dashed") # identity line
```

In some more detail, the gray-dashed identity lines in Figure\ \@ref(fig:w-shapes) imply a linear weighting pattern.
That is, since the images of $w$, $w(p)$, are equal to the respective cumulative probabilities $p$, all cumulative decision weights $\pi$ that are derived according to (12) would be equal to the differences between the respective objective cumulative probabilities and therefore also to the objective probabilities of the outcomes for which the weights are determined.
Such a linear weighting pattern is implied by $\gamma = 1$ and $\delta = 1$. 
Now, both deviations from $\gamma = 1$ and $\delta = 1$ produce a nonlinear graphical shape.
That is, for $\gamma > 1$, the graphs take a S-shape---running below (above) the identity line for small (large) cumulative probabilities---which is more accentuated for larger deviations from $\gamma = 1$.
Importantly, the same *small* numerical difference between two cumulative probabilities $p$ then translates to varying differences between the respective images $w(p)$ depending on the interval on the cumulative probability scale. 
Specifically, because of the S-shaped curvature, the differences between the images $w(p)$ may be smaller than those between the respective cumulative probabilities in the lower and upper part of the cumulative probability scale, but may be considerably greater in the middle part.
As a consequence, the cumulative decisions weights are smaller than the respective objective probabilities for small-probability outcomes of either a low rank or a high rank, but they are larger for small-probability outcomes of a middle rank.^[<!--Open footnote-->Note that the distinction between outcomes of either low/high rank or middle rank is unnecessary for two-outcome prospects [see @tverskyAdvancesProspectTheory1992]. However, for finite many outcomes, the distinction is critical. More so, the fact that CPT may weigh outcomes of the same small probability differently, adds another reasons why the description-experience gap should not be equated with a reversed weighting pattern in CPT [see @hertwigConstructBehaviorGap2018]. I.e., neither could the gap then be interpreted in terms of an as-if underweighting (overweighting) of rare outcomes in DfE (DfD), nor would any of the potential contributors to the gap considered so far [see, e.g., @hertwigDescriptionexperienceGapRisky2009; @wulffMetaanalyticReviewTwo2018] treat some small-probability outcomes differently than other small-probability outcomes.<!--Close footnote-->]
Note that for all $\gamma < 1$ the graph of $w$ takes an inverse S-shape and the entire weighting pattern is reversed.
Moreover, the shape of the weighting function, and thus the weighting pattern, depends strongly on $\delta$, which affects the overall elevation of the graph.
That is, for $\delta < 1$, the interval on the cumulative probability scale over which the weighting function runs below the identity line is greater than the interval over which the weighting function runs above the identity line; and this pattern reverses for $\delta > 1$.
More specifically, $\delta$ shifts across the entire cumulative probability scale the intervals within which small differences between objective cumulative probabilities translate to either smaller or greater cumulative decision weights.
In other words then, for any decrease (increase) in $\delta$, the underweighting of small probability outcomes extends to more outcomes at the lower (higher) end of the outcome range.

@@

To demonstrate the impact of different sampling strategies on the inferences about the latent properties of prospects and the resulting choices in DfE, a simulation study was conducted.
Therefore, computational models were implemented that generate and integrate samples from prospects according to varying parameters reflecting different sampling strategies.

@@

We test a set of 2-prospect choice problems, where one of the prospects contains a safe outcome, i.e., $p(\omega) = 1$ and the other two outcomes where all $p(\omega) \neq 1$.
Both outcomes and probabilities are drawn from uniform distributions, ranging from 0 to 20 for outcomes and from .01 to .99 for probabilities of the smaller outcome of the risky prospect.
To omit dominant prospects within a choice problem, outcomes of the safe prospect always fall between both outcomes of the risky prospect.
Table A1 in contains the test set of 60 choice problems, which were sampled from an initial set of 10,000.
Sampling of gambles was stratified, randomly drawing an equal number of 20 gambles with no, an attractive, and an unattractive rare outcome.
Risky outcomes are considered *"rare"* if their probability is $p < .2$ and *"attractive"* (*"unattractive"*) if they are higher (lower) than the safe outcome.

For each parameter combination of the generating model, 100 synthetic agents are presented with 60 choices problems.
In sum, 100 (parameter combinations) x 100 (agents) x 60 (choices) = 600,000 choices are simulated.

@@

The switching probability $s$ is the probability with which agents draw a single sample from the prospect they did not get their most recent single sample from.
$s$ is varied between .1 to 1 in increments of .1.
The two boundary parameters resemble the concept of a decision threshold, i.e., if a prospect reaches a boundary, it is chosen by the synthetic agent.
The boundary type is either the minimum number of comparisons any prospect must win (absolute boundary) or the minimum difference between the number of won comparisons (relative boundary).
The boundary value $a$ is varied between 1 to 5 in increments of 1.

@@

(ref:fr-rates-p) Effect of Switching Probability on Rates of False Risky and False Safe Choices for 2-Outcome Prospects with High Rank Outcomes of Different Probability

```{r fig.cap="(ref:fr-rates-p)", fig.height=7, fig.width=10}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices_p <- read_csv("data/choices/choices.csv", col_types = cols) # piecewise integration

# false response (fr) rates

## determine normative choice according to latent and sampled EV
fr_rates_p <- choices_p %>%
  mutate(norm = case_when(ev_ratio > 1 ~ "A", ev_ratio < 1 ~ "B"), # normative choice latent EV
         ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2),
         norm_exp = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) # normative choice sampled EV

# fr rates latent EV
fr_rates_p_o <- fr_rates_p %>%
  filter(!is.na(norm)) %>% # exclude choices under EV equality
  group_by(s, boundary, a, rare, norm, choice) %>% # to distinguish type of rare event and type of false response
  summarise(n = n()) %>% # determine absolute number of responses per category
  mutate(rate = round(n/sum(n), 2), # determine correct and false response rates
         type = case_when(norm == "A" & choice == "B" ~ "Safe", norm == "B" & choice == "A" ~ "Risky")) %>% # type false response
  ungroup() %>%
  filter(!is.na(type)) %>% # remove correct response rates
  select(-c(norm, choice, n)) %>%
  mutate(norm_ev = "Latent")

# fr rates sampled EV
fr_rates_p_e <- fr_rates_p %>%
  filter(!is.na(norm_exp)) %>%
  group_by(s, boundary, a, rare, norm_exp, choice) %>%
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2),
         type = case_when(norm_exp == "A" & choice == "B" ~ "Safe", norm_exp == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>%
  select(-c(norm_exp, choice, n)) %>%
  mutate(norm_ev = "Sample")

# merge data frame
fr_rates_p <- bind_rows(fr_rates_p_o, fr_rates_p_e) %>%
  mutate(rare = case_when(rare == "none" ~ "\\in \\[.2,.8\\]",
                          rare == "attractive" ~ "\\in (0,.2)",
                          rare == "unattractive" ~ "\\in (.8,1)")) # Change fact labels

# labeller function for LaTeX math expressions in facet labels
labeller_a <- function(string) {
  TeX(paste("$\\a=$", string, sep = ""))
}

labeller_rare <- function(string) {
  TeX(paste("$\\p_{High}$", string, sep = ""))
}

ggplot(fr_rates_p, aes(s, rate)) +
  facet_grid(rare~a, labeller = labeller(rare = as_labeller(labeller_rare, default = label_parsed), a = as_labeller(labeller_a, default = label_parsed))) +
  labs(x = "s",
       y = "False Response Rate",
       color = "Normative\nEV",
       shape = "False\nResponse") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  labs(x = "Switching Probability",
       y = "False Response Rate",
       color = "Normative\nExpected\nValue",
       shape = "False\nResponse\nType") +
  geom_point(aes(color = norm_ev, shape = type)) +
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis_d(option = "plasma", end = .7) +
  theme_apa()
```

<!-- Test set-->

@@

```{r fig.cap="(ref:weighting-parameters)", fig.height=7, fig.width=10}

# load data

cols <- list(.default = col_double(),
             boundary = col_factor(),
             a = col_factor(),
             parameter = col_factor())
cpt_p <- read_csv("data/estimates/estimates_cpt.csv", col_types = cols) # piecewise integration
cpt_p <- cpt_p %>% mutate(boundary = if_else(boundary == "absolute", "Absolute", "Relative"))

gamma <- cpt_p %>%
  filter(parameter == "gamma") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  labs(x = "Switching Probability",
       y = expression(gamma),
       color = "Switching\nProbability",
       shape = "Boundary") +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(labeller_a, default = label_parsed))) + 
  theme_apa() + 
  theme(axis.title.x = element_blank())

>>>>

bibliography      : references_sampling-in-dfe.bib
csl               : /Users/Linus Hof/Documents/Psychologie/ARC/sampling-in-dfe/apa.csl
appendix          : manuscript_appendix.Rmd

@@

classoption       : "man"
documentclass     : "apa7"
output: papaja::apa6_pdf

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, #omit messages
                      fig.align = "l", #align figures to left margin
                      fig.pos = "t"
                      )
```

```{r packages, include=FALSE}
#load required packages
pacman::p_load(papaja,
               readr,
               tidyr, 
               dplyr,
               ggplot2,
               viridis,
               ggpubr,
               latex2exp)
```

@@

Similar to other behavioral phenomena that could be explained by assuming that the mind is sensitive to the composition of the samples on which it operates [see @fiedlerBewareSamplesCognitiveecological2000], the as-if underweighting of rare outcomes in DfE could be explained by assuming that the mind is sensitive to the information provided by the small samples that people tend to rely on [@hertwigDecisionsExperienceEffect2004; @plonskyRelianceSmallSamples2015; @rakowBiasedSamplesNot2008; @wulffMetaanalyticReviewTwo2018].
That is, following from the laws of large numbers [LLN, see, e.g., @kolmogorovFoundationsTheoryProbability1950], the relative frequency with which a prospect's outcome occurs in a sample of infinite size converges almost surely to its objective probability.
In such a case, then, according to the de Moivre-Laplace theorem, the binomial distribution associated with each prospect's outcome can be approximated by a symmetric normal distribution [cf. @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015].
However, for small samples, the binomial distribution associated with small-probability outcomes is positive skewed, therefore causing the relative frequencies with which these outcomes occur in the sample to be smaller rather than larger than their latent objective probabilities.
Thus, small-probability outcomes tend to be underrepresented in small samples and might not occur at all.

Consistent with this statistical regularity, sequential sampling models of decision making [e.g., @bhatiaSequentialSamplingParadoxes2014; @busemeyerDecisionFieldTheory1993; @markantModelingChoiceSearch2015; @ratcliffComparisonSequentialSampling2004] assume that the information-processing chain upfront choices in DfE takes the form of an evidence accumulation process over time, with choices becoming more accurate the more evidence must be accumulated.
Specifically, in one paradigm used to study DfE, namely, the sampling paradigm [@weberPredictingRiskSensitivity2004], before making a choice between prospects with latent properties, people can, as much as they want, sequentially sample the prospects' outcomes according to their objective probabilities.
To relate the sampling process to the choice, sequential sampling models assume that the sampled outcomes are sequentially accumulated into one or more dynamic decision variables, and that a choice is made once a decision variable exceeds a threshold in favor for one of the prospects.
Accordingly, the number of outcomes required to sample decreases with thresholds, which could account for an as-if underweighting of rare outcomes [see @markantModelingChoiceSearch2015].

However, some studies found the as-if underweighting of rare outcomes in DfE to appear, even if the overall number of sampled outcomes is large and the relative frequencies across all sampled outcomes closely resemble the latent objective probabilities [@hauDescriptionexperienceGapRisky2008; @hauDecisionsExperienceStatistical2010; @ungemachAreProbabilitiesOverweighted2009; see also @wulffMetaanalyticReviewTwo2018].
While this finding demonstrates that the as-if underweighting of rare outcomes in DfE cannot be reduced to a small overall number of sampled outcomes, it may still be the case that the respective deviations from the principle of mathematical expectation were caused by the reliance on small samples.
For example, given a large overall number of sampled outcomes, an as-if underweighting of rare outcomes could be explained, if only a small subset of all the sampled outcomes was considered for the assessment of prospects.
<!---->

Yet, in another approach, @hillsInformationSearchDecisions2010 propose a dependency between sampling patterns and decision strategies.
Considering the choice between two prospects, this approach distinguishes sampling patterns along the frequency of switches between the prospects during the sampling phase and assumes that only outcomes that were sampled from the same prospect between switches are integrated into the same sample. 
Then, it is assumed that each time after a new sample for both prospects was generated, their means are compared, and the choice is determined by the number of won comparisons.^[<!--Open footnote-->Note that @hillsInformationSearchDecisions2010 describe only the two paradigmatic cases, where prospects are either switched just once (comprehensive sampling) or after each sampled outcome (piecewise sampling). Accordingly, in the former case choices depend on a single comparison between the means of samples that contain all outcomes sampled for the respective prospect (summary strategy); in the latter case, however, choices depend on multiple comparisons between samples that contain only one sampled outcome (round-wise strategy). Nevertheless, Hills and Hertwig [-@hillsInformationSearchDecisions2010, p. 1788] acknowledge that "many [sampling] strategies will fall on the continuum between [these cases]", which is why the general description is used above.<!--Close footnote-->]
Since each comparison is weighted equally, such a link between sampling patterns and decision strategies can explain an as-if underweighting of rare outcomes despite a large overall number of sampled outcomes being considered.
More specifically, while the probability of a rare outcome to be sampled increases with the overall number of sampled outcomes, an increasing switching frequency causes the rare outcome not to be considered in an increasing number of comparisons.

@@

@@

\noindent
Given that the value of each cumulative decision weight depends on the two transformed cumulative probabilities in (12)---i.e., the probability of obtaining a positive (negative) outcome equal to or greater (smaller) than a respective outcome $x$, and the probability of obtaining a strictly greater (smaller) outcome---the remaining consequences for the weighting pattern are reviewed by considering the actual weighting function.
Therefore, Figure\ \@ref(fig:weighting-function) illustrates some of the possible graphical shapes of the two-parameter weighting function of @goldsteinExpressionTheoryPreference1987, which, however, is just one of several parameterizations that have been proposed [e.g., @prelecProbabilityWeightingFunction1998; @tverskyAdvancesProspectTheory1992; see @stottCumulativeProspectTheory2006, for an overview].
Now, let each $p$ on the abscissa be one of the four cumulative probabilities from (12). 
Then each graph in Figure\ \@ref(fig:weighting-function) displays the graphical shape of the weighting function
$$
\begin{aligned}
  w : [0,1] &\to [0,1] \\
  p &\mapsto \frac{\delta \times p^{\gamma}}
  {\delta \times p^{\gamma} + (1-p)^{\gamma}}
  \; ,
\end{aligned}
\tag{11}
$$
for the respective values of the parameters $\gamma \in [0,2]$ and $\delta > 0$.
Evidently, both parameters have distinct effects on the graphs' shape, with $\gamma$ affecting the curvature and $\delta$ the elevation.
As a consequence from (12) then, each combination of parameters implies a particular weighting pattern, with some of them being similar and others being rather distinct.

(ref:weighting-function) Possible Graphical Shapes of Goldstein and Einhorn's -@goldsteinExpressionTheoryPreference1987 Weighting Function

```{r weighting-function, fig.cap="(ref:weighting-function)", fig.height=7, fig.width=10}
#compute images of weighting function (wf)
wf <- tibble(p = seq(0, 1, .01)) %>%  #cumulative probabilities
  expand_grid(gamma = seq(.1, 2, .1), #gamma values
              delta = c(.1, .5, 1, 2, 5, 10)) %>% #delta values
  mutate(wp = (delta*(p^gamma))/((delta*p^gamma)+(1-p)^gamma)) #images of wf

#labeller function for facet labels with LateX math expressions 
label_delta <- function(string) {
  TeX(paste("$\\delta=$", string, sep = ""))  
}

#plot shapes of weighting function
wf %>% 
  ggplot(aes(p, wp, group = gamma)) +
  facet_wrap(~delta, labeller = as_labeller(label_delta, default = label_parsed)) + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x = expression(p),
       y = expression(w(p)), 
       color = expression(gamma)) + 
  theme_apa() + 
  geom_line(aes(color = gamma)) +
  scale_color_viridis(option = "plasma") + 
  geom_abline(intercept = 0, slope = 1, color = "gray", linetype = "dashed")
```

In some more detail, the gray-dashed identity lines in Figure\ \@ref(fig:weighting-function) imply a linear weighting pattern.
That is, since the images of $w$, $w(p)$, are equal to the respective cumulative probabilities $p$, all cumulative decision weights $\pi$ that are derived according to (12) would be equal to the differences between the respective objective cumulative probabilities and therefore also to the objective probabilities of the outcomes for which the weights are determined.
Such a linear weighting pattern is implied by $\gamma = 1$ and $\delta = 1$. 
Now, both deviations from $\gamma = 1$ and $\delta = 1$ produce a nonlinear graphical shape.
That is, for $\gamma > 1$, the graphs take a S-shape---running below (above) the identity line for small (large) cumulative probabilities---which is more accentuated for larger deviations from $\gamma = 1$.
Importantly, the same *small* numerical difference between two cumulative probabilities $p$ then translates to varying differences between the respective images $w(p)$ depending on the interval on the cumulative probability scale. 
Specifically, because of the S-shaped curvature, the differences between the images $w(p)$ may be smaller than those between the respective cumulative probabilities in the lower and upper part of the cumulative probability scale, but may be considerably greater in the middle part.
As a consequence, the cumulative decisions weights are smaller than the respective objective probabilities for small-probability outcomes of either a low rank or a high rank, but they are larger for small-probability outcomes of a middle rank.^[<!--Open footnote-->Note that the distinction between outcomes of either low/high rank or middle rank is unnecessary for two-outcome prospects [see @tverskyAdvancesProspectTheory1992]. However, for finite many outcomes, the distinction is critical. More so, the fact that CPT may weigh outcomes of the same small probability differently, adds another reasons why the description-experience gap should not be equated with a reversed weighting pattern in CPT [see @hertwigConstructBehaviorGap2018]. I.e., neither could the gap then be interpreted in terms of an as-if underweighting (overweighting) of rare outcomes in DfE (DfD), nor would any of the potential contributors to the gap considered so far [see, e.g., @hertwigDescriptionexperienceGapRisky2009; @wulffMetaanalyticReviewTwo2018] treat some small-probability outcomes differently than other small-probability outcomes.<!--Close footnote-->]
Note that for all $\gamma < 1$ the graph of $w$ takes an inverse S-shape and the entire weighting pattern is reversed.
Moreover, the shape of the weighting function, and thus the weighting pattern, depends strongly on $\delta$, which affects the overall elevation of the graph.
That is, for $\delta < 1$, the interval on the cumulative probability scale over which the weighting function runs below the identity line is greater than the interval over which the weighting function runs above the identity line; and this pattern reverses for $\delta > 1$.
More specifically, $\delta$ shifts across the entire cumulative probability scale the intervals within which small differences between objective cumulative probabilities translate to either smaller or greater cumulative decision weights.
In other words then, for any decrease (increase) in $\delta$, the underweighting of small probability outcomes extends to more outcomes at the lower (higher) end of the outcome range.

@@

@@

The test set of $60$ choice problems, each consisting of a safe prospect and a two-outcome prospect, was obtained by stratified sampling from an initial set of $10,000$ choice problems.
The stratification was used to assure that some of the choice problems contain a small (high) outcome of a probability smaller than $.20$.
The exact procedure was as follows:
For each of the $10.000$ problems in the initial set, three outcomes were randomly drawn from a uniform distribution over the interval $[0, 20]$, with the smallest and highest of these three outcomes being assigned to the risky prospect in order to omit the case of dominant prospects.
The probability for the smaller outcome of the risky prospect, $p_{small}$, was drawn from a uniform distribution over the interval $[.01, .99]$, with the probability of the higher outcome being set to $p_{high} = 1-p_{small}$.
To obtain the $60$ choice problems of the test set, the initial set was divided into three subsets, where each subset contained either all problems with $p_{high} \in (0,.2)$ or $p_{high} \in [.2,.8]$ or $p_{high} \in (.8,1)$.
From each of these subset, then, $20$ choice problems were randomly sampled.

@@

The evidence accumulation model presented in this paper was implemented as a computational model to simulate the sampling process and the final choice for each of the $60$ choice problems in the test set.
All of the following parameter values were combined with each other:

* **Switching probability**: 
The switching probability $s$, defined as the probability of sampling outcomes from different prospects in direct succession, was varied in the interval $[.1, 1]$ in increments of $.1$.

* **Decision variable**: The decision variable of a prospect either counts the absolute number of won comparisons or how much more comparisons were won relative to the other prospect.

* **Threshold**: 
The value of the threshold $\theta$, determining how much (more) comparisons a prospect must win, was varied in the interval $[1,5]$ in increments of 1.

\noindent
In sum, 100 (parameter combinations) x 60 (choices) x 100 (agents) = 600,000 sampling processes and choices were simulated.

@@

```{r data, include=FALSE}
#load choice data
cols_choices <- list(.default = col_double(),
                     boundary = col_factor(),
                     gamble = col_factor(),
                     rare = col_factor(),
                     agent = col_factor(),
                     choice = col_factor())
choices <- read_csv("data/choices/choices.csv", col_types = cols_choices)

#load CPT estimates
cols_cpt <- list(.default = col_double(),
                 boundary = col_factor(),
                 a = col_factor(),
                 parameter = col_factor())
cpt <- read_csv("data/estimates/estimates_cpt.csv", col_types = cols_cpt)
```

<!-- Plots choice rates -->

(ref:choice-rates) Effect of Switching Probability on Rates of False Risky and False Safe Choices for 2-Outcome Prospects with High Rank Outcomes of Different Probability

```{r choice-rates, fig.cap="(ref:choice-rates)", fig.height=7, fig.width=10}
#prepare data

##determine normative choice according to latent and sampled EV
fr_rates <- choices %>%
  mutate(norm = case_when(ev_ratio > 1 ~ "A", ev_ratio < 1 ~ "B"), #normative choice latent EV
         ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2),
         norm_exp = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) #normative choice sampled EV

##compute false response rates

###latent EV
fr_rates_l <- fr_rates %>%
  filter(!is.na(norm)) %>% #exclude trials with equal EV
  group_by(s, boundary, a, rare, norm, choice) %>% #separate by model parameter and type of rare event
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2), #compute response rates
         type = case_when(norm == "A" & choice == "B" ~ "Safe", #determine false response type
                          norm == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>% #remove correct responses
  select(-c(norm, choice, n)) %>%
  mutate(norm_ev = "Latent") #assign facet label

###sampled EV
fr_rates_s <- fr_rates %>%
  filter(!is.na(norm_exp)) %>%
  group_by(s, boundary, a, rare, norm_exp, choice) %>%
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2),
         type = case_when(norm_exp == "A" & choice == "B" ~ "Safe",
                          norm_exp == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>%
  select(-c(norm_exp, choice, n)) %>%
  mutate(norm_ev = "Sample")

### merge latent and sampled data
fr_rates <- bind_rows(fr_rates_l, fr_rates_s) %>%
  mutate(rare = case_when(rare == "none" ~ "\\in \\[.2,.8\\]", #change facet labels for rare events
                          rare == "attractive" ~ "\\in (0,.2)",
                          rare == "unattractive" ~ "\\in (.8,1)"))

# labeller functions
label_theta <- function(string) {
  TeX(paste("$\\theta=$", string, sep = "")) #threshold parameter theta
}

label_rare <- function(string) {
  TeX(paste("$\\p_{High}$", string, sep = "")) #type of rare event
}

#plot
fr_rates %>%
  ggplot(aes(s, rate)) +
  facet_grid(rare~a, labeller = labeller(rare = as_labeller(label_rare, default = label_parsed), 
                                         a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = "Switching Probability",
       y = "False Response Rate",
       color = "Normative\nExpected\nValue",
       shape = "False\nResponse\nType") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  geom_point(aes(color = norm_ev, shape = type)) +
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis_d(option = "plasma", end = .7) +
  theme_apa()
```

<!-- Plots CPT -->

@@

```{r weighting-parameters, fig.cap="(ref:weighting-parameters)", fig.height=7, fig.width=10}
#prepare data
cpt <- cpt %>% mutate(boundary = if_else(boundary == "absolute", "Absolute", "Relative"))

#plot estimates

##gamma
gamma <- cpt %>%
  filter(parameter == "gamma") %>%
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(label_theta, default = label_parsed))) +
  labs(x = element_blank(), #omit axis title
       y = expression(gamma),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis(option = "plasma") +
  theme_apa()

##delta
delta <- cpt %>%
  filter(parameter == "delta") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(delta),
       color = "Switching\nProbability",
       shape = "Boundary") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  theme_apa() + 
  theme(strip.text.x = element_blank()) #omit facet labels

##merge figures
ggarrange(gamma, delta, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

(ref:value-parameters) Parameter Estimates of the Value Function of @tverskyAdvancesProspectTheory1992 and of the Logit Choice Rule for Different Simulation Parameters
